---
title: "Exploratory HR Analytics"
author: "Daniel El Sabbagh Velasco"
date: "20 de noviembre de 2017"
output: html_document
---
#LOADING DATASET

```{r, message=FALSE, warning=FALSE}
library(readr)
HR_comma_sep <- read_csv("C:/Users/danie/Desktop/HUMAN RESOURCES ANALYTICS - EQUIPO ANALISIS DE DATOS/HR_comma_sep.csv")
View(HR_comma_sep)
```
```{r}
head(HR_comma_sep)
```


# MISSING DATA
## We check the amount of missing data from our dataset

```{r}
which(is.na(HR_comma_sep))
```

## We can see that we don't have missing data in our dataset. 
# CREATING NEW VARIABLES
## It is interesting create a new variable that relates number_project and time_spend_company
```{r, message=FALSE}
library(dplyr)
HR_dataset <- mutate(HR_comma_sep, project_per_year = number_project/time_spend_company)
View(HR_dataset)
```

Now we can see the head of our dataset with the new variable: project_per_year
```{r}
head(HR_dataset)
```

# EXPLORING THE DATA

We have around 15.000 employees in our company, and we know about them: satisfaction_level, last_evaluation, number_project, average_monthly_hours, time_spend_company, work_accident, left, promotion_last_5years, sales and salary. 
We make a exploratory table: 

```{r}
summary(HR_dataset)
```

We can see that the mean of satisfaction_level is around 0.61, or the performance average is around 0.71. Also the average_monthly_hours is around 200, time_spend_company is around 3.498 years on average and project_per_year is around 1.21 on average. Besides around 14,4% of our workers have a work and 23,81% of our workers left the work and only the 0.021% of the workers had a promotion on the last five years. 

## NORMALITY OF VARIABLES
We check the normality of variables with two graphs: histograms and box-plot. 

```{r}
attach(HR_dataset)
par(mfrow=c(1,2))
hist(satisfaction_level, col = "red", xlim = c(0,1))
boxplot(satisfaction_level, col = "red", ylim= c(0,1), main = "Boxplot of Satisfaction_level")
hist(last_evaluation, col = "purple", xlim = c(0,1))
boxplot(last_evaluation, col = "purple", ylim= c(0,1), main = "Boxplot of Last_evaluation")
hist(number_project, col="green", breaks = 7)
boxplot(number_project, col="green", main = "Boxplot of number_project")
hist(average_montly_hours, col = "pink", xlim= c(50, 350))
boxplot(average_montly_hours, col = "pink", main = "Boxplot of Average_monthly_hours")
hist(time_spend_company, col = "blue", breaks = 8)
boxplot(time_spend_company, col = "blue", main = "Boxplot of time_spend_company")
hist(project_per_year, col = "orange", xlim= c(0,3))
boxplot(project_per_year, col = "orange", main = "Boxplot of project_per_year")

```

As we can see, there are only two variables that are'n normal: time_spend_company and project_per_year. 
Besides, we have 4 factors in our dataset with we can realise different graphs with two or more factors, as corrplots or barplots. 
We make a corrplot and a scatterplot in order to see the relation between our variables. To do this, we need change our categorical variables into numeric variables. 


```{r}
require(car)
HR_dataset$salarynum<-recode(HR_dataset$salary, "'low'=1")
HR_dataset$salarynum<-recode(HR_dataset$salarynum, "'medium'=2")
HR_dataset$salarynum<-recode(HR_dataset$salarynum, "'high'=3")
HR_dataset$salarynum<-as.numeric(HR_dataset$salarynum)

library(dplyr)
library(graphics)
library(corrplot)
HR_dataset_corrplot <- select(HR_dataset, -sales, -salary)
HR_correlation <- cor(HR_dataset_corrplot)
library(corrplot)
par(mfrow=c(1,2))
corrplot(HR_correlation, method = "color", tl.cex = 0.6)
corrplot(HR_correlation, method = "number", tl.cex = 0.6)
par(mfrow=c(1,1))
pairs(HR_correlation)
```

## WHO LEAVES THE COMPANY

We can ask ourselves, ''who are leaving the company?''. To answer this question, we can take only the leavers of the company and do barplots. 
```{r, message=FALSE}
library(ggplot2)
HR_leavers <- subset(HR_dataset, HR_dataset$left == 1)
attach(HR_leavers)
```


```{r}
par(mfrow=c(1,1))
hist(satisfaction_level, col = "blue", xlim = c(0,1))
hist(last_evaluation, col = "blue")
hist(number_project, col= "blue")
hist(average_montly_hours, col= "blue")
hist(time_spend_company, col= "blue")
hist(Work_accident, col = "blue")

HR_dataset$salary<-factor(HR_dataset$salary,levels=c("low","medium","high")) 
ggplot(HR_dataset,aes(x=salary, fill=factor(left)))+
  geom_bar(stat='count',position='dodge')
```

## PRINCIPAL COMPONENT ANALYSIS

We make a principal component analysis to see the relation between individuals and variables in only two or three dimensions. 
```{r}
HR_dataset_pca <- princomp(HR_correlation)
summary(HR_dataset_pca)
plot(HR_dataset_pca)
```

As we can see, we can explain the 68% of the variance with 3 principal components. Nevertheless we can take only 2 PCs in order to see easier than 3 PCs. 

```{r}
HR_dataset_pca$loadings
```

We can see there the relevance of the variables to build the PCs. We can see that the second PC is built by satisfaction_level, number_project, etc., and the first PC is built by time_spend_company or project_per_year, for example. 
Besides, we make a pca biplot in order to see how the individuals are related between them and which variables are characterizing them. 

```{r, message = FALSE}
library(ggbiplot)
```

```{r}
ggbiplot(HR_dataset_pca)
```

