---
title: "Biplot"
author: "Roberto Díaz"
date: "16 de diciembre de 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(readr)
HR_comma_sep_v2 <- read_csv("~/Human_Resources_Analytics-Kaggle_DS/final_files/databases/HR_comma_sep_v2.csv")
```

```{r include=FALSE}
install.packages("http://biplot.usal.es/classicalbiplot/multbiplot-in-r/multbiplotrtar.gz", repos = NULL, type="source")
```

We need the package "MultBiplotR"
```{r warning=FALSE, message=FALSE}
require(MultBiplotR)
```

We need all our variables to be cuantitative to be able to do a biplot.

```{r warning=FALSE, message=FALSE}
require(dplyr)
cuanti<-select(HR_comma_sep_v2, -c(department, salary))
```

##RECODE SATISFACTION LEVEL

We are going to recode the variable satisfaction level to a qualitative variable see the different levels if different colours in our plot. We are going to do the same in our plot with the salary.

```{r warning=FALSE, message=FALSE}
require(car)
cuanti$satisnom<-recode(cuanti$satisfaction_level, "0.01:0.25='VL'")
cuanti$satisnom<-recode(cuanti$satisnom, "0.26:0.50='L'")
cuanti$satisnom<-recode(cuanti$satisnom, "0.51:0.75='H'")
cuanti$satisnom<-recode(cuanti$satisnom, "0.76:1='VH'")
cuanti$satisnom<-as.factor(cuanti$satisnom)
```

#BIPLOT


```{r}
biplot<-PCA.Biplot(scale(cuanti[1:9]))
plot(biplot)
```
In this plot we can barely do some interpretation (besides the relationship between variables), so we are going to delete the Labels and put some colour.
```{r warning=FALSE, message=FALSE}
biplot=AddCluster2Biplot(biplot, ClusterType="us", Groups=cuanti$satisnom)
plot(biplot, LabelInd = FALSE, xlim=c(-4,4), ylim=c(-4,4), CexInd=0, PlotClus = TRUE, TypeClus = "ch", ClustCenters = T)
```
```{r}
plot(biplot, LabelInd = FALSE, CexInd=0.5, ColorInd=cuanti$satisnom)
legend(x="topright", c("VL", "L", "H", "VH"), pch=1, col=c("Purple", "Green", "Red", "Skyblue"))
```
We can see here two plots, one with clusters and another with all the individuals.

##Salary

Now we are going to see the same with salary. 
```{r}
HR_comma_sep_v2$salary<-as.factor(HR_comma_sep_v2$salary)
```

```{r}
biplot=AddCluster2Biplot(biplot, ClusterType="us", Groups=HR_comma_sep_v2$salary)
plot(biplot, LabelInd = FALSE, xlim=c(-4,4), ylim=c(-4,4), CexInd=0, PlotClus = TRUE, TypeClus = "ch", ClustCenters = T)
```
```{r}
plot(biplot, LabelInd = FALSE, CexInd=0.5, ColorInd=HR_comma_sep_v2$salary)
legend(x="topright", c("LOW", "MEDIUM", "HIGH"), pch=1, col=c("Green", "Blue", "Red"))
```


#HJ BIPLOT

An HJ-Biplot has a good representation on both variables and individuals, but it is supposed that it doesn´t represent well the original data.
```{r}
hjbiplot<-HJ.Biplot(scale(cuanti[1:9]))
plot(hjbiplot)
```
We can see that the vectors of the variables are too long compared to the dispersion of the individuals, so we have to zoom to see the situation of them. Doing that, we cannot see which vector belongs to which variable, so we have to remember the first plot.
```{r}
plot(hjbiplot, xlim=c(-4,4), ylim=c(-4,4))
```

```{r}
hjbiplot=AddCluster2Biplot(hjbiplot, ClusterType="us", Groups=cuanti$satisnom)
plot(hjbiplot, LabelInd = FALSE, xlim=c(-4,4), ylim=c(-4,4), CexInd=0, PlotClus = TRUE, TypeClus = "ch", ClustCenters = T)
```
```{r}
plot(hjbiplot, LabelInd = FALSE, CexInd=0.5, xlim=c(-4,4), ylim=c(-4,4), ColorInd=cuanti$satisnom)
legend(x="topright", c("VL", "L", "H", "VH"), pch=1, col=c("Purple", "Green", "Red", "Skyblue"))
```

##Salary


```{r}
hjbiplot=AddCluster2Biplot(hjbiplot, ClusterType="us", Groups=HR_comma_sep_v2$salary)
plot(hjbiplot, LabelInd = FALSE, xlim=c(-4,4), ylim=c(-4,4), CexInd=0, PlotClus = TRUE, TypeClus = "ch", ClustCenters = T)
```
```{r}
plot(hjbiplot, LabelInd = FALSE, CexInd=0.5, xlim=c(-4,4), ylim=c(-4,4), ColorInd=HR_comma_sep_v2$salary)
legend(x="topright", c("LOW", "MEDIUM", "HIGH"), pch=1, col=c("Green", "Blue", "Red"))
```
